MyTitle

### 2024-11-08 paper/main.tex
updated

By choosing $\mathbb{P}$ appropriately, we can construct $M[G]$ with various properties.
What holds or does not hold in $M[G]$ can be identified using the forcing relation.
The forcing relation is defined recursively on formulas in a forcing language.
The forcing language is an extension of tha language of set theory by adding the elements of $M^{\mathbb{P}}$ as constants.
\begin{dfn} \label{dfn:forcing_relation}
  We define the \emph{forcing relation} $\Vdash$ for formulas in the forcing language and $p \in \mathbb{P}$ inductively
  \footnote{
    Specifically, the forcing relation is first defined for atomic formulas $\dot{x} = \dot{y}$ and $\dot{x} \in \dot{y}$ by mutual recursion,
    inductively on the $\mathbb{P}$-rank of $\dot{x}$ and $\dot{y}$.
    Then, the definition is extended to all formulas in the forcing language by induction on the complexity of formulas.
  }
  as follows:
  \begin{enumerate}
    \item $p \Vdash \dot{x} = \dot{y} \Leftrightarrow \forall \dot{z} \in \mathrm{dom}(\dot{x}) \cup \mathrm{dom}(\dot{y})\,
\forall q \leq_{\mathbb{P}} p \, (q \Vdash \dot{z} \in \dot{x} \leftrightarrow q \Vdash \dot{z} \in \dot{y})$
    \item $p \Vdash \dot{x} \in \dot{y} \Leftrightarrow \forall q \leq_{\mathbb{P}} p\, \exists r \leq_{\mathbb{P}} q \, 
    \exists s \in \mathbb{P} \, \exists \dot{z} \in M^{\mathbb{P}} \, (\langle \dot{z}, s \rangle \in \dot{y} \land r \leq_{\mathbb{P}} s \land r \Vdash \dot{x} = \dot{z} )$
    \item $p \Vdash \phi \land \psi \Leftrightarrow p \Vdash \phi \land p \Vdash \psi$
    \item $p \Vdash \neg \phi \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \neg (q \Vdash \phi)$
    \item $p \Vdash \exists x \phi(x) \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \exists r \leq_{\mathbb{P}} q \, \exists \dot{x} \in M^{\mathbb{P}} \, (r \Vdash \phi(\dot{x}))$
  \end{enumerate}
\end{dfn}

@@

\begin{cor}
  Let $p \in \mathbb{P}$, $\varphi$ be a formula and $\dot{x} \in M^{\mathbb{P}}$, then
  $$p \Vdash \varphi(\dot{x}) \Leftrightarrow \text{for any generic filter } G \text{ containing } p, M[G] \vDash \varphi(\dot{x}^G)$$
\end{cor}

@@

Paulson and Gunther et al. declared locales to handle internal models of ZF set theory in their work \cite{paulson_AC_consistency} \cite{gunther_forcing}.
Here, we introduce Gunther's locale \texttt{"M_ctm"}.

\subsubsection{The locale \texttt{"forcing\_data"} by Gunther et al.}

@@

\subsubsection{The Relativized Forcing Relation $\Vdash_{\mathrm{HS}}$}
Instead, we formalize the relativized forcing relation $\Vdash_{\mathrm{HS}}$ and prove that symmetric extensions are models of ZF using it.
This relation is defined by replacing all quantifiers restricted to $M^{\mathbb{P}}$
in the definition of the forcing relation with quantifiers restricted to $\mathrm{HS}$.
By modifying the formalization of forcing by Gunther et al. \cite{gunther_forcing}
\footnote{For non-atomic formulas, the definition of the forcing relation by Gunther et al. \cite{gunther_forcing} 
actually differs from Definition \ref{dfn:forcing_relation}.
In their definition, the unique logical connective is NAND, and the unique quantifier is $\forall$.
Additionally, in the definition, quantifiers restricted to $M$ appear instead of those restricted to $M^{\mathbb{P}}$.}
, we were able to define this relation and prove its properties relatively easily.

>>>>

By choosing $\mathbb{P}$ appropriately, we can construct $M[G]$ with various properties.
What holds or does not hold in $M[G]$ can be identified using the forcing relation.
The forcing relation is defined recursively on formulas in a forcing language.
The forcing language is an extension of the language of set theory by adding the elements of $M^{\mathbb{P}}$ as constants.
\begin{dfn} \label{dfn:forcing_relation}
  We define the \emph{forcing relation} $\Vdash$ for formulas in the forcing language and $p \in \mathbb{P}$ inductively
  \footnote{
    Specifically, the forcing relation is first defined for atomic formulas $\dot{x} = \dot{y}$ and $\dot{x} \in \dot{y}$ by mutual recursion,
    inductively on the $\mathbb{P}$-rank of $\dot{x}$ and $\dot{y}$.
    Then, the definition is extended to all formulas in the forcing language by induction on the complexity of formulas.
  }
  as follows:
  \begin{enumerate}
    \item $p \Vdash \dot{x} = \dot{y} \Leftrightarrow \forall \dot{z} \in \mathrm{dom}(\dot{x}) \cup \mathrm{dom}(\dot{y})\,
\forall q \leq_{\mathbb{P}} p \, (q \Vdash \dot{z} \in \dot{x} \leftrightarrow q \Vdash \dot{z} \in \dot{y})$
    \item $p \Vdash \dot{x} \in \dot{y} \Leftrightarrow \forall q \leq_{\mathbb{P}} p\, \exists r \leq_{\mathbb{P}} q \, 
    \exists s \in \mathbb{P} \, \exists \dot{z} \in M^{\mathbb{P}} \, (\langle \dot{z}, s \rangle \in \dot{y} \land r \leq_{\mathbb{P}} s \land r \Vdash \dot{x} = \dot{z} )$
    \item $p \Vdash \phi \land \psi \Leftrightarrow p \Vdash \phi \land p \Vdash \psi$
    \item $p \Vdash \neg \phi \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \neg (q \Vdash \phi)$
    \item $p \Vdash \exists x \phi(x) \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \exists r \leq_{\mathbb{P}} q \, \exists \dot{x} \in M^{\mathbb{P}} \, (r \Vdash \phi(\dot{x}))$
  \end{enumerate}
\end{dfn}
We say that $p$ \emph{forces} $\phi$ if $p \Vdash \phi$.

@@

\begin{cor}\label{cor:truth_lemma_cor}
  Let $p \in \mathbb{P}$, $\varphi$ be a formula and $\dot{x} \in M^{\mathbb{P}}$, then
  $$p \Vdash \varphi(\dot{x}) \Leftrightarrow \text{for any generic filter } G \text{ containing } p, M[G] \vDash \varphi(\dot{x}^G)$$
\end{cor}

@@

Paulson and Gunther et al. declared locales to handle internal models of ZF set theory in their work \cite{paulson_AC_consistency, gunther_forcing}, respectively.
For example, the locale \texttt{"M\_ctm"} by Gunther et al. fixes \texttt{"M"} and \texttt{"enum"},
and introduces assumptions that \texttt{"M"} holds the axioms of ZF set theory, 
\texttt{"M"} is a transitive set, and \texttt{"enum"} is a enumeration of \texttt{"M"}.
Therefore, we can work in the context of a c.t.m.\  of ZF by working within this locale.

\subsubsection{Forcing}
Gunther et al. formalized forcing method in ZF set theory \cite{gunther_forcing}.
Their formalization follows Kunen's book \cite{kunen2011}, which uses the c.t.m.\  approach.

They defined the notion of forcing as the locale \texttt{"forcing\_notion"} 
with the constants \texttt{"P"}, \texttt{"leq"}, and \texttt{"one"}.
\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
locale forcing_notion =
    fixes P leq one
    assumes one_in_P:     "one ∈ P"
      and leq_preord:       "preorder_on(P,leq)"
      and one_max:          "∀p∈P. ⟨p,one⟩∈leq"
\end{minted}
\end{isaframe}

They also declared the locale \texttt{"forcing\_data"} 
by combining the locale \texttt{"forcing\_notion"} with the locale \texttt{"M\_ctm"} and assume some additional properties. 
\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
locale forcing_data = forcing_notion + M_ctm +
    assumes P_in_M:         "P ∈ M"
      and leq_in_M:         "leq ∈ M"
\end{minted}
\end{isaframe}

Within this locale, The predicate \texttt{"p ⊩ φ env"} is defined.
This represents that \texttt{"p"} forces the formula \texttt{"φ"} under the environment \texttt{"env"}.
Additionally, the following
\footnote{
In Definition\ref{dfn:forcing_relation},
we explained that the forcing relation is defined for the formulas in the forcing language.
According to that definition, \texttt{"env"} should be a list of elements of $M^{\mathbb{P}}$;
however, in the following lemmas, \texttt{"env"} is a list of elements of $M$.
In fact, the definition of the forcing relation by Gunther et al. differs from Definition \ref{dfn:forcing_relation}.
They defined the forcing relation not for the formulas in the forcing language 
but for the formulas in the extended language of set theory by adding all elements of $M$ as constants.
Additionally, only NAND is used as the logical connective, and only $\forall$ is used as the quantifier of the formulas.
} is their formalization of Theorem \ref{thm:truth_lemma} and Corollary \ref{cor:truth_lemma_cor}.
\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
lemma truth_lemma:
    assumes 
        "φ∈formula" "M_generic(G)"
    shows 
        "⋀env. env∈list(M) ⟹ arity(φ)≤length(env) ⟹ 
          (∃p∈G. p ⊩ φ env)   ⟷   M[G], map(val(G),env) ⊨ φ"

lemma definition_of_forcing:
    assumes
        "p∈P" "φ∈formula" "env∈list(M)" "arity(φ)≤length(env)"
    shows
        "(p ⊩ φ env) ⟷
        (∀G. M_generic(G) ∧ p∈G  ⟶  M[G], map(val(G),env) ⊨ φ)"
\end{minted}
\end{isaframe}
Where \texttt{"M\_generic(G)"} represents that \texttt{"G"} is a generic filter on \texttt{"P"} in \texttt{"M"},
\texttt{"val(G)"} is the interpretation by \texttt{"G"},
and \texttt{"map"} is the map function on lists.

@@

\subsubsection{The Relativized Forcing Relation $\Vdash_{\mathrm{HS}}$}
Instead, we formalize the relativized forcing relation $\Vdash_{\mathrm{HS}}$ and prove that symmetric extensions are models of ZF using it.
By modifying the formalization of forcing by Gunther et al. \cite{gunther_forcing}, 
we were able to define this relation and prove its properties relatively easily.
